\pdftitle

### 2018-05-13 maxNt_v2.tex
V2

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-13T22:11:43+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

Maximum-entropy models
\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} have been used in
neuroscience with this kind of purpose and for diverse brain areas at least
since the 1990s
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
Part of this literature focused on the range of applicability of \me\
distributions
\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
We assume that our readers are at least vaguely familiar with such models.
In brief, they build a probability distribution for neuronal activity using
only data about simultaneous firings of two neurons, or of three, and so
on. Maximum-entropy probability distributions can be introduced and justified in
three main ways:
\begin{enumerate}
\item from the maximum-entropy method
  \cite{jaynes1957,jaynes1963,sivia1996_r2006};
\item as asymptotic distributions of Bayesian models called
  \enquote{exchangeable} with entropic priors
  \cites{jaynes1986d_r1996,skilling1989b,skilling1990,rodriguez1989,rodriguez2002,portamana2017}[see
  also][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
{jaynes1979b};
\item as asymptotic distributions of Bayesian models based on sufficient
  statistics
  \cite{jaynes1982,%[\sect~I]
jaynes1982b,portamana2017}.
\end{enumerate}

*** usually \enquote{explain} has a probabilistic meaning. We want to see
whether information about moments give reasonable predictions about the
rest. Necessity of making this precise. Three ways:

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-13T23:58:06+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

Maximum-entropy models
\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984} have been used in
neuroscience with this kind of purpose and for diverse brain areas at least
since the 1990s
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015}.
Part of this literature focused on the range of applicability of \me\
distributions
\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.
We assume that our readers are at least vaguely familiar with such models.
In brief, they build a probability distribution for neuronal activity using
only data about simultaneous firings of two neurons, or of three, and so
on. Let us take the case of the first two moments just as an example. The
corresponding maximum-entropy probability distributions can be derived in
three main ways:
\begin{enumerate}
\item From the maximum-entropy method
  \cite{jaynes1957,jaynes1963,sivia1996_r2006}. The distribution is chosen
  as the broadest one, in the Shannon-entropy sense, among those having
  particular first and second moments.
\item As asymptotic distributions of Bayesian models called
  \enquote{exchangeable} with entropic priors
  \cites{jaynes1986d_r1996,skilling1989b,skilling1990,rodriguez1989,rodriguez2002,portamana2017}[see
  also][\chap~VI]{schroedinger1946_r1948}%[pp.~52--55]
  {jaynes1979b}. The distribution is the limit, when the number of time
  bins is very large, of
  \begin{equation}
    \label{eq:predictive_exchangeability}
    \p(\text{\small activity at some time bin} \|
    \text{\small 1st \amp\ 2nd empirical moments from observed bins},\text{\small other info});
  \end{equation}
  obtained from Bayes's theorem \emph{given} the observed empirical moments
  and a particular initial joint distribution.
\item As asymptotic distributions of Bayesian models based on sufficient
  statistics
  \cite{jaynes1982,%[\sect~I]
jaynes1982b,portamana2017}. The distribution is the limit, when the number of time
bins is very large, of
\begin{equation}
  \label{eq:general_prediction}
  \p(\text{\small activity at one time bin} \|
  \text{\small activities at the remaining bins},\text{\small other info})
\end{equation}
under the assumption that only the values of the first and second empirical
moments of the activities are \emph{relevant} for the probability above, \ie\
\begin{multline}
  \label{eq:informal_sufficiency}
  \p(\text{\small activity at one time bin} \|
  \text{\small activities at the other bins},\text{\small other info}) ={}\\
  \p(\text{\small activity at some time bin} \|
  \text{\small 1st \amp\ 2nd empirical moments},\text{\small other info});
\end{multline}
in other words, the assumption that the first and second moments are
\emph{sufficient statistics}. This limit generally does not depend on the initial
joint distribution.
\end{enumerate}

usually \enquote{explain} has a probabilistic meaning. We want to see
whether information about moments give reasonable predictions about the
rest. Necessity of making this precise. Three ways:
